---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# frequentdirections [![Build Status](https://travis-ci.com/shinichi-takayanagi/frequentdirections.svg?branch=master)](https://travis-ci.com/shinichi-takayanagi/frequentdirections) 

Implementation of Frequent-Directions algorithm for efficient matrix sketching [E. Liberty, SIGKDD2013]


## Installation

```{r gh-installation, eval = FALSE}
# Not yet onCRAN
install.packages("frequentdirections")

# Or the development version from GitHub:
install.packages("devtools")
devtools::install_github("shinichi-takayanagi/frequentdirections")
```

## Example

### Download example data

Here, we use [Handwritten digits USPS dataset](https://www.kaggle.com/bistaumanga/usps-dataset/version/1) as sample data.
In the following example, we assume that you save the above sample data into `/tmp` directory.

### Load data

The dataset has 7291 train and 2007 test images in `h5` format. 
The images are 16*16 grayscale pixels.
```{r load-data}
library("h5")
file <- h5file("/tmp/usps.h5")
x <- file["train/data"][]
y <- file["train/target"][]
str(x)
```

### Plot example image
Example the number `8`
```{r plot-example-image}
image(matrix(x[338,], nrow=16, byrow = FALSE))
```

### Plot SVD
Plot the original data on the first and second singular vector plane.
```{r plot-svd}
x <- scale(x)
frequentdirections::plot_svd(x, y)
```

### Matrix Sketching
#### l = 8 case
```{r frequentdirections-8}
eps <- 10^(-8)
# 7291 x 256 -> 8 * 256 matrix
b <- frequentdirections::sketching(x, 8, eps)
frequentdirections::plot_svd(x, y, b)
```

#### l = 32 case
```{r frequentdirections-32}
# 7291 x 256 -> 32 * 256 matrix
b <- frequentdirections::sketching(x, 32, eps)
frequentdirections::plot_svd(x, y, b)
```

#### l = 128 case
```{r frequentdirections-128}
# 7291 x 256 -> 128 * 256 matrix
b <- frequentdirections::sketching(x, 128, eps)
frequentdirections::plot_svd(x, y, b)
```

This result is almost the same with the original data SVD expression.

That's why we can think that the original data is expressed with only `128` rows.
